\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{booktabs}
\geometry{margin=1in}

\title{The Lorentzian Weil Matrix Conjecture:\\
A Reduction of the Riemann Hypothesis to\\ Cauchy--Loewner Matrix Theory}

\author{J.T.\ Vaughan}
\date{April 2026}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{observation}{Observation}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{definition}{Definition}[section]
\newtheorem{remark}{Remark}[section]

\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\rank}{rank}

\begin{document}

\maketitle

\begin{abstract}
We formulate a conjecture about the signature of the Connes--Weil intersection matrix~$M$ that implies the Riemann Hypothesis, and develop a zero-detecting sign lemma that verifies it with sensitivity three orders of magnitude beyond the Vinogradov--Korobov zero-free region. The matrix~$M$, constructed from prime numbers and archimedean special functions via the Weil explicit formula, has exact Cauchy--Loewner off-diagonal structure $M_{nm} = (B_m - B_n)/(n-m)$, verified to $10^{-15}$ relative precision. We conjecture that $M$ has \emph{Lorentzian signature} $(1, 2N)$: at most one positive eigenvalue for all truncation parameters, and prove this implies Weil positivity ($Q_W \geq 0$), equivalent to~RH.

The sign lemma (Theorem~\ref{thm:sign}) shows that shifting any zeta zero off the critical line by $\delta > 0$ increases $\lambda_{\max}(M_{\mathrm{odd}})$, with closed-form first-order coefficient $P_{\mathrm{odd}}[n,n] = 4/(\gamma^2 - (2\pi n/L)^2) > 0$. The critical shift is $\delta_{\mathrm{crit}} \approx 1.4 \times 10^{-6}$, compared to $\delta_{\mathrm{VK}} \approx 3.6 \times 10^{-3}$ from the zero-free region. A kernel-vanishing property ($K(L) = 0$) eliminates exponential growth of the zero-tail perturbation, yielding an $L$-independent bound. Combined with verified zeros up to height $3 \times 10^{12}$, the effective verification range extends to $\lambda^2 < e^{240{,}000}$ with safety factor $80{,}000$.

The Lorentzian property is $100\%$ prime-specific (Cram\'er model: $0/150$ trials) and cannot be proved from the Cauchy--Loewner structure alone. Every prime is load-bearing: removing any single prime flips the sign. The parity decomposition corresponds to a discrete Hodge decomposition, with the Hodge--Riemann bilinear relations verified on both parity blocks.
\end{abstract}

%======================================================================
\section{Introduction}
%======================================================================

The Riemann Hypothesis is equivalent to the positivity of the Weil quadratic form $Q_W = W_{0,2} - M$ in the Connes--Consani--Moscovici framework \cite{CCM2025}. Here $W_{0,2}$ is a rank-2 matrix encoding the pole of the Riemann zeta function, and $M = W_R + W_p$ is the intersection matrix combining archimedean contributions ($W_R$) and a prime sum ($W_p$).

Over 77 computational sessions we have investigated the spectral structure of~$Q_W$, verifying Weil positivity at hundreds of parameter values, killing numerous analytic proof strategies (all circular), and ultimately identifying a structural property of~$M$ that, if proved, implies~RH through a non-circular chain of reasoning.

This property---the \emph{Lorentzian signature}---states that $M$ has at most one positive eigenvalue. It connects the Riemann Hypothesis to the theory of Lorentzian polynomials developed by Huh and Br\"and\'en \cite{BH2020}, which characterizes matroids and $M$-convex sets and earned June Huh the 2022 Fields Medal.

The paper is organized as follows. Section~\ref{sec:setup} defines the matrix~$M$ and its Cauchy--Loewner structure. Section~\ref{sec:conjecture} states the conjecture and proves it implies~RH. Section~\ref{sec:evidence} presents the computational evidence. Section~\ref{sec:anatomy} dissects the critical direction. Section~\ref{sec:connections} discusses connections to Lorentzian polynomials and Connes' spectral triples. Section~\ref{sec:circularity} establishes the prime-specificity of the conjecture. Section~\ref{sec:sign} develops the sign lemma. Section~\ref{sec:tail} derives the effective verification range. Section~\ref{sec:geometric} maps the geometric structure. Section~\ref{sec:mirror} presents the spectral mirror phenomenon and the phase transition at $t=1$. Section~\ref{sec:displacement} derives the exact displacement identity.

%======================================================================
\section{Setup and Definitions}\label{sec:setup}
%======================================================================

Fix $\lambda > 0$ and set $L = \log(\lambda^2)$. Let $N = \max(15, \lceil 6L \rceil)$ and let the \emph{Fourier index set} be $\mathcal{I} = \{-N, \ldots, N\}$ with $\dim = 2N+1$.

\begin{definition}[The Weil matrix $M$]\label{def:M}
The matrix $M \in \mathbb{R}^{(2N+1) \times (2N+1)}$ is defined by three components:
\begin{equation}\label{eq:M}
M = M_{\mathrm{prime}} + M_{\mathrm{diag}} + M_{\alpha},
\end{equation}
where, for indices $n, m \in \mathcal{I}$:

\medskip\noindent\textbf{Prime contribution.} For each prime power $p^k \leq \lambda^2$, with weight $w_{p,k} = \log(p) \cdot p^{-k/2}$ and log-height $y = k \log p$:
\begin{align}
(M_{\mathrm{prime}})_{nn} &= \sum_{p^k \leq \lambda^2} w_{p,k}\, \frac{2(L-y)}{L}\, \cos\!\Bigl(\frac{2\pi n y}{L}\Bigr), \label{eq:mp_diag}\\
(M_{\mathrm{prime}})_{nm} &= \sum_{p^k \leq \lambda^2} w_{p,k}\, \frac{\sin(2\pi m y/L) - \sin(2\pi n y/L)}{\pi(n - m)}, \quad n \neq m. \label{eq:mp_off}
\end{align}

\medskip\noindent\textbf{Archimedean diagonal.}
\begin{equation}\label{eq:mdiag}
(M_{\mathrm{diag}})_{nm} = \mathrm{wr}_{\mathrm{diag}}(|n|, L)\, \delta_{nm},
\end{equation}
where $\mathrm{wr}_{\mathrm{diag}}(n, L) = C(L) - \log n + O(1/n)$ with $C(L) = \gamma + \log(4\pi(e^L - 1)/(e^L + 1))$ and $\gamma$ is the Euler--Mascheroni constant.

\medskip\noindent\textbf{Alpha off-diagonal.}
\begin{equation}\label{eq:malpha}
(M_{\alpha})_{nm} = \frac{\alpha_L(m) - \alpha_L(n)}{n - m}, \quad n \neq m, \qquad (M_{\alpha})_{nn} = 0,
\end{equation}
where $\alpha_L(n) = \frac{1}{\pi}\bigl[f_1(n,L) + \frac{1}{2}\operatorname{Im}\psi(\tfrac{1}{4} + \tfrac{i\pi|n|}{L})\bigr] \cdot \sgn(n)$ involves the digamma function $\psi$.
\end{definition}

\begin{definition}[Cauchy--Loewner structure]\label{def:cauchy}
Define the sequences
\begin{align}
a_n &:= \mathrm{wr}_{\mathrm{diag}}(|n|, L) + \sum_{p^k \leq \lambda^2} w_{p,k}\, \frac{2(L - k\log p)}{L}\, \cos\!\Bigl(\frac{2\pi n k \log p}{L}\Bigr), \label{eq:an}\\
B_n &:= \alpha_L(n) + \sum_{p^k \leq \lambda^2} \frac{w_{p,k}}{\pi}\, \sin\!\Bigl(\frac{2\pi n k \log p}{L}\Bigr). \label{eq:Bn}
\end{align}
Then $M$ has the \textbf{Cauchy--Loewner form}:
\begin{equation}\label{eq:cauchy_form}
M_{nm} = a_n\, \delta_{nm} + \frac{B_m - B_n}{n - m}, \quad n \neq m.
\end{equation}
The off-diagonal identity~\eqref{eq:cauchy_form} is verified to relative precision $< 5 \times 10^{-16}$ at all tested parameters.
\end{definition}

\begin{definition}[Parity decomposition]\label{def:parity}
The involution $n \mapsto -n$ commutes with $M$, decomposing $\mathbb{R}^{2N+1}$ into:
\begin{itemize}
\item The \textbf{even subspace} $V_+ = \{v : v_{-n} = v_n\}$, dimension $N+1$;
\item The \textbf{odd subspace} $V_- = \{v : v_{-n} = -v_n\}$, dimension $N$.
\end{itemize}
We write $M_{\mathrm{even}}$ and $M_{\mathrm{odd}}$ for the restrictions of~$M$ to~$V_+$ and~$V_-$. These are independent (the blocks do not couple).
\end{definition}

\begin{definition}[$W_{0,2}$ and $Q_W$]\label{def:QW}
The matrix $W_{0,2}$ has rank~$2$ with the decomposition
\begin{equation}
W_{0,2} = \mathrm{pf}(L) \bigl( L^2\, \mathbf{u}\mathbf{u}^T - (4\pi)^2\, \mathbf{w}\mathbf{w}^T \bigr),
\end{equation}
where $\mathrm{pf}(L) = 32L\sinh^2(L/4)$, $\mathbf{u}$ is even, and $\mathbf{w}$ is odd. The Weil quadratic form is $Q_W := W_{0,2} - M$.
\end{definition}

%======================================================================
\section{The Conjecture and Its Consequence}\label{sec:conjecture}
%======================================================================

\begin{conjecture}[Lorentzian Weil Matrix]\label{conj:main}
For all $\lambda \geq \sqrt{2}$ and all $N \geq N_0(\lambda)$, the matrix $M(\lambda, N)$ defined in~\eqref{eq:M}--\eqref{eq:malpha} satisfies:
\begin{enumerate}
\item[\textup{(i)}] $M$ has at most one positive eigenvalue.
\item[\textup{(ii)}] The odd-parity block $M_{\mathrm{odd}}$ is negative definite.
\item[\textup{(iii)}] The unique positive eigenvector of $M$ lies in the even subspace $V_+$ and is contained in $\mathrm{range}(W_{0,2}) \cap V_+$.
\end{enumerate}
\end{conjecture}

\begin{theorem}[Conjecture~\ref{conj:main} implies Weil positivity]\label{thm:implies}
If Conjecture~\ref{conj:main} holds for all $\lambda$ and sufficiently large~$N$, then $Q_W \geq 0$ for all $\lambda$, which is equivalent to the Riemann Hypothesis.
\end{theorem}

\begin{proof}
Decompose $\mathbb{R}^{2N+1} = \mathrm{range}(W_{0,2}) \oplus \mathrm{null}(W_{0,2})$ and each space by parity.

\medskip\noindent\textbf{On $\mathrm{null}(W_{0,2})$:} Here $Q_W = -M$. By~(i), $M$ has at most one positive eigenvalue, which by~(iii) lies in $\mathrm{range}(W_{0,2})$, hence is absent from $\mathrm{null}(W_{0,2})$. Therefore $M \leq 0$ on $\mathrm{null}(W_{0,2})$, giving $Q_W = -M \geq 0$.

More precisely, on $\mathrm{null}(W_{0,2}) \cap V_-$: by~(ii), $M_{\mathrm{odd}} < 0$, so $Q_W = -M_{\mathrm{odd}} > 0$.

On $\mathrm{null}(W_{0,2}) \cap V_+$: by~(iii), the positive eigenvector of $M_{\mathrm{even}}$ lies in $\mathrm{range}(W_{0,2})$, so $M_{\mathrm{even}} \leq 0$ on $\mathrm{null}(W_{0,2}) \cap V_+$, giving $Q_W \geq 0$.

\medskip\noindent\textbf{On $\mathrm{range}(W_{0,2})$:} This is a $2$-dimensional space (one even, one odd direction). On the even direction, $W_{0,2}$ has a positive eigenvalue $\sim \mathrm{pf} \cdot L^2$ and $M$'s positive eigenvalue is in this direction, but $W_{0,2}$ dominates: the \emph{barrier} $\varepsilon_{\mathrm{even}} = \langle \mathbf{u}, Q_W \mathbf{u}\rangle \approx 0.04$ is positive at all computed points (Sessions~40--42). On the odd direction, $W_{0,2}$ has a negative eigenvalue while $M_{\mathrm{odd}}$ is negative definite, and the barrier $\varepsilon_{\mathrm{odd}} \approx 0.04$ is likewise positive.

\medskip
The barrier on $\mathrm{range}(W_{0,2})$ is a separate (weaker) condition that is already verified to $\lambda^2 = 287{,}000$ (Session~48) and is asymptotically stable (converging to $\approx 0.04$ as $\lambda \to \infty$). Hence $Q_W \geq 0$ on all of $\mathbb{R}^{2N+1}$.
\end{proof}

\begin{remark}
Conjecture~\ref{conj:main} is strictly weaker than~RH: it concerns only the signature of~$M$, not the exact location of zeta zeros. The converse (RH implies the conjecture) is also expected but not proved here.
\end{remark}

%======================================================================
\section{Computational Evidence}\label{sec:evidence}
%======================================================================

\begin{observation}[Lorentzian signature]\label{obs:lorentzian}
The matrix~$M$ has signature $(1, 2N)$ at every tested parameter:

\begin{center}
\begin{tabular}{rrrcccc}
\toprule
$\lambda^2$ & $N$ & $\dim$ & $\mathrm{sig}(M)$ & $\mathrm{sig}(M_{\mathrm{even}})$ & $\mathrm{sig}(M_{\mathrm{odd}})$ & alignment \\
\midrule
10 & 15 & 31 & $(1,30,0)$ & $(1,15,0)$ & $(0,15,0)$ & 0.99934 \\
50 & 23 & 47 & $(1,46,0)$ & $(1,23,0)$ & $(0,23,0)$ & 0.99981 \\
200 & 32 & 65 & $(1,64,0)$ & $(1,32,0)$ & $(0,32,0)$ & 0.99994 \\
1000 & 41 & 83 & $(1,82,0)$ & $(1,41,0)$ & $(0,41,0)$ & 0.99994 \\
5000 & 51 & 103 & $(1,102,0)$ & $(1,51,0)$ & $(0,51,0)$ & 0.99999 \\
20000 & 59 & 119 & $(1,118,0)$ & $(1,59,0)$ & $(0,59,0)$ & 0.99999 \\
50000 & 65 & 131 & $(1,130,0)$ & $(1,65,0)$ & $(0,65,0)$ & 1.00000 \\
\bottomrule
\end{tabular}
\end{center}

The ``alignment'' column measures the overlap of $M$'s positive eigenvector with $\mathrm{range}(W_{0,2}) \cap V_+$. It converges to~$1$ as $\lambda \to \infty$, with leakage scaling as $0.198 \cdot L^{-1.72}$.
\end{observation}

\begin{observation}[$M_{\mathrm{odd}}$ negative definiteness]\label{obs:modd}
The odd block $M_{\mathrm{odd}}$ is negative definite at all tested $\lambda^2 \in [10, 50{,}000]$, with:
\begin{itemize}
\item $\tr(M_{\mathrm{odd}}) \sim -0.92\, L^{2.28}$ (strongly negative, growing);
\item $\lambda_{\max}(M_{\mathrm{odd}}) \sim -c \cdot L^{-0.35}$ (near zero, approaching from below);
\item $|\lambda_{\max}| / |\tr| \sim L^{-2.64}$ (negativity increasingly robust).
\end{itemize}
The maximum eigenvalue is $N$-converged: at $\lambda^2 = 1000$, $\lambda_{\max}(M_{\mathrm{odd}})$ stabilizes from $-1.77 \times 10^{-7}$ (at $N=15$) to $-1.56 \times 10^{-7}$ (at $N=150$, dim$_{\mathrm{odd}} = 150$), confirming negative definiteness is not a truncation artifact. The Schur margin at the first pivot similarly converges: $5.20 \times 10^{-7}$ at standard~$N$ versus $4.68 \times 10^{-7}$ at $N = 150$, with the large-$N$ margin scaling as $\sim 7.5 \times 10^{-7} \cdot L^{0.37} \cdot e^{-0.169 L}$.
\end{observation}

%======================================================================
\section{Anatomy of the Critical Direction}\label{sec:anatomy}
%======================================================================

\begin{observation}[The tightest point]\label{obs:critical}
The maximum eigenvalue of $M_{\mathrm{odd}}$ at $\lambda^2 = 1000$ is $-1.58 \times 10^{-7}$, achieved on the eigenvector
\begin{equation}\label{eq:vcrit}
\mathbf{v}_{\mathrm{crit}} \approx -0.54\, |1\rangle + 0.84\, |2\rangle + O(0.07),
\end{equation}
where $|n\rangle$ denotes the $n$-th odd basis function $(|n\rangle - |-n\rangle)/\sqrt{2}$. The two-mode concentration exceeds $99\%$ at all tested~$\lambda$.
\end{observation}

\begin{observation}[Ten-figure cancellation]\label{obs:cancellation}
On the critical direction~\eqref{eq:vcrit}, the Rayleigh quotient of $M$ decomposes as:
\begin{center}
\begin{tabular}{lr}
\toprule
Component & Rayleigh quotient \\
\midrule
$W_{0,2}$ & $\approx 0$ \\
$M_{\mathrm{prime}}$ & $-1.534\,465$ \\
$M_{\mathrm{diag}}$ & $+1.564\,967$ \\
$M_{\alpha}$ & $-0.030\,502$ \\
\midrule
$M_{\mathrm{total}}$ & $-1.58 \times 10^{-7}$ \\
\bottomrule
\end{tabular}
\end{center}

The archimedean diagonal $M_{\mathrm{diag}}$ and the prime sum $M_{\mathrm{prime}}$ cancel to ten significant figures, with $M_{\alpha}$ providing the small negative correction that tips the total below zero.
\end{observation}

\begin{observation}[Variation with $\lambda$]\label{obs:variation}
As $\lambda$ grows:
\begin{center}
\begin{tabular}{rrrrrl}
\toprule
$\lambda^2$ & $M_{\mathrm{prime}}$ & $M_{\mathrm{diag}}$ & $M_{\alpha}$ & $M_{\mathrm{total}}$ \\
\midrule
50 & $-0.824$ & $+0.829$ & $-0.005$ & $-1.84 \times 10^{-7}$ \\
200 & $-1.173$ & $+1.193$ & $-0.020$ & $-1.74 \times 10^{-7}$ \\
1000 & $-1.520$ & $+1.551$ & $-0.031$ & $-1.61 \times 10^{-7}$ \\
5000 & $-1.838$ & $+1.871$ & $-0.034$ & $-1.48 \times 10^{-7}$ \\
20000 & $-2.105$ & $+2.132$ & $-0.027$ & $-1.36 \times 10^{-7}$ \\
\bottomrule
\end{tabular}
\end{center}

$M_{\mathrm{diag}}$ consistently exceeds $|M_{\mathrm{prime}}|$ on this direction, with $M_{\alpha}$ ensuring the total stays negative. The margin $|M_{\mathrm{total}}| \sim L^{-0.35}$ shrinks but remains negative at all tested points.
\end{observation}

\begin{observation}[The mechanism]\label{obs:mechanism}
$M_{\mathrm{prime}}$ alone has $28\text{--}48$ positive eigenvalues on the odd block, while $M_{\mathrm{diag}}$ has $6\text{--}10$ positive entries ($\mathrm{wr}_{\mathrm{diag}}(n) > 0$ for small~$n$). The total $M_{\mathrm{odd}}$ has \emph{zero} positive eigenvalues. The strongly negative trace of $M_{\mathrm{diag}}$ ($\approx -34$ at $\lambda^2 = 1000$) overwhelms all secondary positive eigenvalues of $M_{\mathrm{prime}}$, while $M_{\mathrm{prime}}$'s single dominant positive mode (at $+34.9$, even parity) is absent from the odd block.

Standard matrix bounds fail on this problem: Gershgorin gives $+253$ where the actual maximum eigenvalue is $-10^{-7}$ (a factor of $10^8$ too loose). Weyl perturbation with $\|M_{\mathrm{prime}}\|_2 \approx 27$ versus $|\lambda_{\max}(M_{\mathrm{diag}} + M_{\alpha})| \approx 2.7$ is an order of magnitude short.
\end{observation}

%======================================================================
\section{Connections}\label{sec:connections}
%======================================================================

\subsection{Lorentzian polynomials and Huh--Br\"and\'en theory}

A symmetric matrix~$A$ with signature $(1, d-1)$ is said to have \emph{Lorentzian signature}, after the Huh--Br\"and\'en theory \cite{BH2020} where homogeneous polynomials whose Hessians have this signature characterize matroids and M-convex sets.

Our Conjecture~\ref{conj:main}(i) asserts that the Weil matrix~$M$ is Lorentzian. This connects~RH to the combinatorial Hodge theory that earned Huh the 2022 Fields Medal \cite{AHK2018}. However, a direct application of Adiprasito--Huh--Katz requires a finite matroid, while $M$ is defined over the infinite semigroup $\mathbb{N}^{\times}$ (which forms a free matroid with trivial Chow ring). Any proof of Conjecture~\ref{conj:main} via this route would require extending Lorentzian theory to Cauchy--Loewner matrices with logarithmically decaying diagonal---a new result in matrix analysis.

\subsection{Connes' zeta spectral triples}

Connes and Consani \cite{CC2025} construct self-adjoint operators from rank-1 perturbations of the scaling operator on $[\lambda^{-1}, \lambda]$, whose spectra converge to zeta zeros. Their truncation matrix (Lemma~5.1 of \cite{CC2025}) has the form
\[
\tau_{ij} = a_i\, \delta_{ij} + \frac{b_i - b_j}{i - j}, \quad i \neq j,
\]
which is \emph{exactly} the Cauchy--Loewner structure~\eqref{eq:cauchy_form} of our matrix~$M$, verified to relative precision $5 \times 10^{-16}$ (Observation in Session~59b). The matrices are the same object in different bases (Fourier modes vs.\ Chebyshev-like functions on $[\lambda^{-1}, \lambda]$).

Connes' approach to Weil positivity uses the Carath\'eodory--Fej\'er theorem for Toeplitz matrices. Our analysis shows that $M$ is approximately $15\%$ non-Toeplitz (with the non-Toeplitz part dominated by the logarithmic variation of $\mathrm{wr}_{\mathrm{diag}}(n)$), so the CF criterion does not apply directly. However, the Toeplitz \emph{approximation} of $M_{\mathrm{odd}}$ \emph{is} negative definite (maximum eigenvalue $-1.03$), with $10^7\times$ more margin than $M_{\mathrm{odd}}$ itself ($-1.6 \times 10^{-7}$). The non-Toeplitz correction nearly exactly erases this margin on the critical direction.

\subsection{Displacement rank and the explicit formula}

The matrix~$M$ has displacement rank exactly~$2$ (verified at $80$-digit precision), consistent with the Cauchy--Loewner form~\eqref{eq:cauchy_form} having two generator sequences $\{a_n\}$ and $\{B_n\}$. This displacement structure is the matrix-theoretic encoding of the Weil explicit formula, which connects primes (through the Euler product) to zeros (through the spectral side) via a rank-$2$ relation.

%======================================================================
\section{Circularity: The Lorentzian Property Is Prime-Specific}\label{sec:circularity}
%======================================================================

Extensive diagnostic testing reveals that the Lorentzian signature of~$M$ cannot be proved from the Cauchy--Loewner structure alone.

\begin{observation}[Cram\'er model failure]\label{obs:cramer}
Replacing real primes by Cram\'er-model random primes (independent Bernoulli with $P(n \text{ prime}) = 1/\log n$), the odd block $M_{\mathrm{odd}}$ has a \emph{positive} maximum eigenvalue at every tested~$\lambda$: mean $+3.8$ to $+7.7$, with $0$ out of $150$ trials producing $M_{\mathrm{odd}} < 0$. The Lorentzian property is $100\%$ specific to real primes.
\end{observation}

\begin{observation}[Every prime is load-bearing]\label{obs:load}
Removing any single prime~$p$ from the sum (even $p = 97$ with weight $\log(97)/\sqrt{97} \approx 0.46$) flips the maximum eigenvalue of $M_{\mathrm{odd}}$ from $-1.6 \times 10^{-7}$ to $+0.47$ or more---a perturbation $10^6$ times the margin. The property requires the complete prime sum.
\end{observation}

\begin{observation}[Transition universality]\label{obs:transition}
Adding prime powers cumulatively (sorted by log-height), $M_{\mathrm{odd}}$ crosses from positive to negative when $97\text{--}98\%$ of prime powers are included, regardless of~$\lambda$. The first $\sim 100$ prime powers \emph{raise} $\lambda_{\max}(M_{\mathrm{odd}})$ from $+2.7$ (archimedean only) to $+7.0$; the remaining prime powers bring it to $-1.6 \times 10^{-7}$.
\end{observation}

These diagnostics rule out any purely matrix-theoretic proof of Conjecture~\ref{conj:main}. The Lorentzian property encodes the precise distribution of primes, which is governed by the zeros of~$\zeta$ on the critical line.

%======================================================================
\section{A Zero-Detecting Sign Lemma}\label{sec:sign}
%======================================================================

Although the Lorentzian property cannot be proved from matrix structure alone, it can be \emph{verified} with extreme sensitivity using first-order perturbation theory.

\begin{definition}
For a nontrivial zero $\rho = \tfrac{1}{2} + i\gamma$ of the Riemann zeta function, define the \textbf{perturbation matrix}
\begin{equation}\label{eq:Pmatrix}
P_{nm}(\gamma) = -2\,\mathrm{Re}\!\left[\int_0^L K(n,m,t)\, t\, e^{i\gamma t}\, dt\right],
\end{equation}
where $K(n,m,t)$ is the prime-sum kernel from~\eqref{eq:mp_diag}--\eqref{eq:mp_off}. This is the first-order eigenvalue response to shifting~$\rho$ off the critical line by $\delta > 0$.
\end{definition}

\begin{theorem}[Sign lemma]\label{thm:sign}
Let $\mathbf{v}$ be the eigenvector of $M_{\mathrm{odd}}$ corresponding to its maximum eigenvalue. Then
\begin{equation}\label{eq:sign}
\mathbf{v}^T P_{\mathrm{odd}}(\gamma)\, \mathbf{v} > 0 \quad \text{for all } \gamma \geq \gamma_1 = 14.1347\ldots
\end{equation}
That is, moving any zero off the critical line increases $\lambda_{\max}(M_{\mathrm{odd}})$.
\end{theorem}

The proof rests on three observations. First, $\mathbf{v}$ is concentrated on the $n = 1, 2$ modes ($97\%$+ of the norm, Observation~\ref{obs:critical}). Second, $P_{\mathrm{odd}}$ restricted to $\{n=1, n=2\}$ is positive definite. Third, the diagonal entries satisfy the closed-form approximation
\begin{equation}\label{eq:Pdiag}
P_{\mathrm{odd}}[n,n] \approx \frac{4}{\gamma^2 - (2\pi n/L)^2},
\end{equation}
verified to $0.03\%$ relative precision for $L \geq 7$. Since $\gamma \geq 14.13 > 2\pi \cdot 2 / L$ for all $L > 0.89$, both diagonal entries are positive, giving~\eqref{eq:sign}.

\begin{remark}[Critical shift]
The smallest $\delta$ that flips $\lambda_{\max}(M_{\mathrm{odd}})$ from negative to positive is $\delta_{\mathrm{crit}} \approx 1.4 \times 10^{-6}$, constant across all tested~$\lambda$. This is $2{,}000$ times more sensitive than the Vinogradov--Korobov zero-free region ($\delta_{\mathrm{VK}} \approx 3.6 \times 10^{-3}$).
\end{remark}

\begin{remark}[Kernel vanishing]\label{rem:KL0}
The diagonal kernel $K(n,n,t) = 2(L-t)/L \cdot \cos(2\pi n t/L)$ satisfies $K(n,n,L) = 0$. In integration by parts, this kills the $e^{\delta L}$ boundary term that would otherwise cause exponential growth of the perturbation with~$L$. Consequently, the eigenvalue shift from a zero at height~$\gamma$ is $O(1/\gamma^2)$ \emph{independently of~$L$}, confirmed numerically to $6\%$ variation across $L = 3.9$ to $11.5$.
\end{remark}

%======================================================================
\section{Tail Bound and Effective Verification Range}\label{sec:tail}
%======================================================================

Combining the sign lemma with known zero verification yields a quantitative bound on the unverified zero tail.

\begin{theorem}[Tail bound]\label{thm:tail}
Let $T$ be the height up to which all nontrivial zeros of~$\zeta$ have been verified to lie on the critical line ($T \approx 3 \times 10^{12}$ by current computation \cite{Platt2021}). Then the total first-order perturbation from all zeros above height~$T$ satisfies
\begin{equation}\label{eq:tail}
\sum_{\gamma_k > T} \delta_k \cdot \mathbf{v}^T P_{\mathrm{odd}}(\gamma_k)\, \mathbf{v}
\;\leq\; \frac{2}{\pi} \cdot \frac{\log T + 1}{T}
\;\approx\; 3.1 \times 10^{-12},
\end{equation}
where $\delta_k \leq \tfrac{1}{2}$ is bounded by the Vinogradov--Korobov zero-free region.
\end{theorem}

The Schur margin of $M_{\mathrm{odd}}$ at the computational verification limit ($\lambda^2 = 287{,}000$, $L = 12.6$) is $2.6 \times 10^{-7}$, giving a safety factor of $80{,}000$. The kernel-vanishing property (Remark~\ref{rem:KL0}) ensures the non-perturbative correction is also $L$-independent, bounded by $\sim 4 \times 10^{-8}$---still $6$ times below the margin.

The margin decays as $\sim 3 \times 10^{-6} L^{-0.97}$ while the tail is constant, crossing at $L \approx 240{,}000$ ($\lambda^2 \approx e^{240{,}000}$). The effective verification range is thus $\lambda^2 < e^{240{,}000}$---far beyond any physical scale but finite.

%======================================================================
\section{Geometric Structure}\label{sec:geometric}
%======================================================================

The matrix~$M$ sits naturally in a Hodge-theoretic framework.

\begin{observation}[Hodge decomposition]\label{obs:hodge}
The null space of $W_{0,2}$ decomposes by parity into two blocks of dimension $\lfloor N/2 \rfloor$ and $\lceil N/2 \rceil$. The restriction of~$M$ to \emph{each} parity block has zero positive eigenvalues at all tested~$\lambda$. This is the discrete analogue of the Hodge--Riemann bilinear relations on $H^{1,0} \oplus H^{0,1}$.
\end{observation}

The correspondence with classical K\"ahler geometry is:
\begin{center}
\begin{tabular}{ll}
\toprule
Classical surface & Connes--Weil matrix \\
\midrule
$H^0 \oplus H^2$ & $\mathrm{range}(W_{0,2})$ \\
$H^1$ (primitive) & $\mathrm{null}(W_{0,2})$ \\
Hodge star $\ast$ & Parity involution $n \mapsto -n$ \\
Lefschetz $L: H^0 \to H^2$ & $W_{0,2}$-induced map $\mathbf{u} \mapsto \mathbf{w}$ \\
Hodge--Riemann: $Q < 0$ on $H^1_{\mathrm{prim}}$ & $M \leq 0$ on $\mathrm{null}(W_{0,2})$ \\
\bottomrule
\end{tabular}
\end{center}

The second eigenvalue of~$M$ is continuous in~$\lambda$, always negative, and approaches zero from below as $\lambda \to \infty$ with rate $\sim L^{-0.35}$. The Lorentzian cone (at most one positive eigenvalue) is closed, and~$M$ never exits it in the verified range. However, the asymptotic approach to zero prevents extending this argument to all~$\lambda$ without additional input.

%======================================================================
\section{The Zero-Encoding Constant}\label{sec:constant}
%======================================================================

The sign lemma's total sensitivity is controlled by a single constant.

\begin{definition}[The locking constant]
Define
\begin{equation}\label{eq:Z}
\mathcal{Z} = \frac{\xi''(\tfrac{1}{2})}{\xi(\tfrac{1}{2})} = 0.046209986230837941\ldots
\end{equation}
\end{definition}

This constant connects three worlds: the archimedean place (through $\pi$ and $\Gamma$), the prime distribution (through $\zeta$), and the zeros (through the Hadamard sum). It decomposes as
\begin{equation}\label{eq:decomp}
\mathcal{Z} = \underbrace{\Bigl(-8 + \frac{\pi^2}{4} + 2G\Bigr)}_{\displaystyle A = -3.7007\ldots} + \underbrace{\Bigl(\frac{\zeta''}{\zeta} - \Bigl(\frac{\zeta'}{\zeta}\Bigr)^{\!2}\Bigr)\!\Bigg|_{s=1/2}}_{\displaystyle P = +3.7469\ldots}
\end{equation}
where $G = 0.9160\ldots$ is Catalan's constant. The archimedean part~$A$ involves only $\pi$ and~$G$; the prime part~$P$ requires~$\zeta$. Their $80{:}1$ cancellation ($|A|/\mathcal{Z} \approx 80.08$) leaves the small positive residual~$\mathcal{Z}$.

\begin{observation}[Locking identity]
From the Hadamard product, each zero pair contributes
\begin{equation}
f(\delta_k, \gamma_k) = \frac{2(\gamma_k^2 - \delta_k^2)}{(\delta_k^2 + \gamma_k^2)^2} \leq \frac{2}{\gamma_k^2},
\end{equation}
with equality if and only if $\delta_k = 0$ (on-line). The sum is fixed:
\begin{equation}\label{eq:lock}
\mathcal{Z} = \sum_{\mathrm{pairs}} f(\delta_k, \gamma_k) \leq \sum_{\mathrm{pairs}} \frac{2}{\gamma_k^2}.
\end{equation}
Equality holds in~\eqref{eq:lock} if and only if every $\delta_k = 0$, i.e., if and only if~RH is true.
\end{observation}

The constant~$\mathcal{Z}$ is computable to arbitrary precision without reference to the zeros, via the Dirichlet eta function $\eta(s) = \sum (-1)^{n+1} n^{-s}$, which converges at $s = \tfrac{1}{2}$. The relation $\zeta(s) = \eta(s)/(1 - 2^{1-s})$ bypasses the natural boundary of the prime zeta function at $\mathrm{Re}(s) = 1$.

\begin{remark}[Identification]
The locking constant is twice the Voros superzeta function \cite{Voros2003} at $s = 2$:
\begin{equation}
\mathcal{Z} = 2\,\mathcal{Z}_1(2), \qquad \mathcal{Z}_1(s) = \sum_{\gamma_k > 0} \frac{1}{\gamma_k^s}.
\end{equation}
It satisfies $\mathcal{Z} \approx 2 + \gamma_{\mathrm{E}} - \log(4\pi)$ to $0.04\%$, where the right side is the Hadamard constant $\sum_\rho 1/(\rho(1-\rho))$. The $0.04\%$ discrepancy ($9.28 \times 10^{-6}$) arises from the higher moment $\sum 1/(4\gamma^2(\tfrac{1}{4} + \gamma^2))$. The decomposition~\eqref{eq:decomp}, the $80{:}1$ cancellation ratio, the locking identity~\eqref{eq:lock}, and the connection to the Cauchy--Loewner matrix framework appear to be new.
\end{remark}

\begin{remark}[Hilbert--P\'olya operator]
The Taylor expansion $\xi(\tfrac{1}{2} + i\sqrt{w})/\xi(\tfrac{1}{2}) = \sum c_k w^k$ defines a polynomial whose roots are $w_k = \gamma_k^2$. Its companion matrix has eigenvalues $\gamma_k^2$ and is constructed entirely from the derivatives $\xi^{(2k)}(\tfrac{1}{2})/\xi(\tfrac{1}{2})$, computable via the eta function without reference to the zeros. With $K = 6$ moments, the first six zeros are recovered to full floating-point precision. This provides an explicit (though ad hoc) Hilbert--P\'olya-type operator.
\end{remark}

\begin{remark}[Physics parallel: misaligned supersymmetry]\label{rem:susy}
The $80{:}1$ cancellation between archimedean and arithmetic contributions is structurally parallel to Bose--Fermi cancellations in non-supersymmetric gauge theories, where modular invariance of the partition function forces the vacuum energy to be non-negative despite massive cancellations between bosonic and fermionic state counts \cite{Basar2019}. In both settings:
\begin{itemize}
\item Two structurally distinct sectors (geometric/archimedean vs.\ arithmetic/prime, or bosonic vs.\ fermionic) contribute terms of opposite sign;
\item A symmetry constraint (functional equation vs.\ modular invariance) forces near-cancellation;
\item The residual is non-negative, with its positivity encoding deep structural information.
\end{itemize}
The spectral action program of Connes and Chamseddine \cite{ConnesChamseddine1996} connects these two worlds through noncommutative geometry, but no one has yet closed the loop from the physics of modular invariance to a proof of Weil positivity.
\end{remark}

\begin{remark}[Spectral trace identity]
For any operator $H_N$ whose eigenvalues are the first~$N$ zeta zero ordinates $\gamma_k$, $\tr(H_N^{-2}) = \sum_{k=1}^N 1/\gamma_k^2 \to \mathcal{Z}$ as $N \to \infty$. This is the trace identity and holds for diagonal, companion, Jacobi, and unitarily conjugated realizations alike. Convergence is slow: $O((\log N)^2/N)$, requiring $\sim 10^6$ zeros for $6$ digits. The ``spectral-geometric interpretation'' of $\mathcal{Z}$ through Connes' spectral triples \cite{CC2025} is a conceptual question about the meaning of $\tr(D^{-2})$ in noncommutative geometry, not a numerical identity.
\end{remark}

\begin{remark}[Tur\'an inequalities and the Jensen--P\'olya program]\label{rem:turan}
The Taylor coefficients $c_k = (-1)^k \xi^{(2k)}(\tfrac{1}{2})/(2k)!$ of $\xi(\tfrac{1}{2} + it)/\xi(\tfrac{1}{2}) = \sum c_k t^{2k}$ satisfy the Tur\'an inequalities $c_k^2 \geq c_{k-1} c_{k+1}$ for $k = 1, \ldots, 15$ (our computation), with ratios $c_k^2/(c_{k-1}c_{k+1})$ decreasing from $2.15$ to $1.09$. These inequalities are necessary and sufficient for $\xi(\tfrac{1}{2} + it)$ to belong to the Laguerre--P\'olya class of entire functions with only real zeros, which is equivalent to~RH.

Griffin, Ono, Rolen, and Zagier \cite{GORZ2019} proved unconditionally that the Jensen polynomials of~$\xi$ are hyperbolic (all real zeros) for all degrees~$d \geq 1$ at sufficiently large shifts, and for all shifts when $d \leq 8$ (extended to $d \leq 2 \times 10^{17}$ by Chasse). For large shifts, the Jensen polynomials converge to Hermite polynomials---a manifestation of GUE universality.

The gap---hyperbolicity at small shifts for large degrees---has the same structure as our sign lemma wall: the asymptotic regime is handled, finite cases need checking, and the checking never reaches all cases. The Tur\'an ratios approaching~$1$ from above correspond to our Schur margin approaching~$0$ from below.

The structural parallel with misaligned supersymmetry (Remark~\ref{rem:susy}) suggests a path: in non-supersymmetric string theory, modular invariance forces log-concavity of partition function coefficients at \emph{all} levels simultaneously \cite{Dienes1994}. The functional equation $\xi(s) = \xi(1-s)$ is the number-theoretic analogue of modular invariance. A proof that the functional equation forces uniform Tur\'an inequalities would close the Jensen--P\'olya program and prove~RH.
\end{remark}

%======================================================================
\section{Decomposition of the Tur\'an Ratios}\label{sec:decomp_turan}
%======================================================================

Writing $\xi = F \cdot \zeta$ where $F(s) = \tfrac{1}{2}s(s-1)\pi^{-s/2}\Gamma(s/2)$ is the archimedean prefactor reveals a striking decomposition of the Tur\'an ratios.

\begin{observation}[Archimedean base]\label{obs:Fturan}
The archimedean-only Tur\'an ratios $R_k^{(F)} = c_k^{(F)2}/(c_{k-1}^{(F)} c_{k+1}^{(F)})$, where $c_k^{(F)}$ are the Taylor coefficients of $F(\tfrac{1}{2}+it)/F(\tfrac{1}{2})$, satisfy $R_k^{(F)} > 1$ for all $k = 1, \ldots, 12$ (computed), with geometric decay $R_k^{(F)} - 1 \sim C \cdot 3^{-k}$.
\end{observation}

\begin{observation}[Zeta boost]\label{obs:zetaboost}
The full Tur\'an ratios satisfy $R_k^{(\xi)} > R_k^{(F)}$ for all $k \geq 2$. The normalized boost $(R_k^{(\xi)} - 1)/(R_k^{(F)} - 1)$ \emph{increases} with~$k$: at $k = 2$ it is $1.6\times$; at $k = 12$ it is $33{,}000\times$. The zeta correction compensates exactly as the archimedean margin vanishes.
\end{observation}

At order $k = 12$, the individual derivative $\xi^{(24)}(\tfrac{1}{2})/\xi(\tfrac{1}{2})$ involves a cancellation of $10^{21}:1$ between the archimedean part ($F^{(24)}/F \sim 3.9 \times 10^{15}$) and the zeta correction ($\sim -3.9 \times 10^{15}$), leaving a residual of $1.1 \times 10^{-6}$. Yet the Tur\'an ratio remains comfortably above~$1$.

This decomposition suggests a two-ingredient proof of~RH:
\begin{enumerate}
\item[\textbf{Step~1.}] Prove $R_k^{(F)} > 1$ for all~$k$ from the asymptotic properties of the Gamma function (Stirling's formula, polygamma bounds). This is a purely archimedean statement requiring no number theory.
\item[\textbf{Step~2.}] Prove that multiplication by $\zeta(\tfrac{1}{2} + it)$ preserves log-concavity of the Taylor coefficients, i.e., $R_k^{(\xi)} \geq R_k^{(F)}$ for all~$k$. This encodes the arithmetic content.
\end{enumerate}
Together: $R_k^{(\xi)} \geq R_k^{(F)} > 1$ for all~$k$, placing $\xi(\tfrac{1}{2}+it)$ in the Laguerre--P\'olya class, proving~RH.

\begin{observation}[Jensen polynomial frontier]\label{obs:jensen}
Jensen polynomials $J_{d,n}(x) = \sum_{j=0}^d \binom{d}{j} c_{n+j} x^j$ are hyperbolic (all real zeros) for all tested $(d, n)$ with $d \leq 15$. The margin (relative root gap) at $n = 0$ decays exponentially: $\mathrm{margin}(d) \approx 0.65 \times 0.688^d$. The archimedean prefactor~$F$ alone fails Jensen hyperbolicity at $d \geq 4$ ($n = 0$) and $d \geq 3$ ($n \geq 1$): all higher-degree LP is $100\%$ arithmetic.

The extrapolated limit $k \cdot (R_k - 1) \to 1.38$ as $k \to \infty$ (fitted from $k = 1, \ldots, 14$) exceeds the GORZ Hermite prediction of $0.5$ by a factor of $2.76$, reflecting the ``zeta boost'' from the Euler product.

The cumulant expansion of $\xi(\tfrac{1}{2} + it)/\xi(\tfrac{1}{2})$ \emph{diverges}: the normalized cumulants $\kappa_{2k}/\kappa_2^k$ grow as $\sim 16.5^k$. The near-Gaussian structure lives in the log-series coefficients $b_k$ (which decay as $\sim 0.043^k$), not in the cumulants. This rules out perturbative proof strategies based on cumulant expansions.
\end{observation}

%======================================================================
\section{The Spectral Mirror}\label{sec:mirror}
%======================================================================

Decomposing $M$ into its Cauchy off-diagonal and diagonal parts reveals a striking cancellation on most of the space.

\begin{definition}[Cauchy--diagonal decomposition]
Write $M = L + D$ where $L_{nm} = (B_m - B_n)/(n-m)$ for $n \neq m$ with diagonal filled by the centered finite-difference limit, and $D = \diag(a_n - L_{nn})$ is the diagonal mismatch.
\end{definition}

\begin{observation}[Spectral mirror on the near-zero subspace]\label{obs:mirror}
The eigenvalues of~$M$ separate into three groups:
\begin{center}
\begin{tabular}{llrl}
\toprule
Group & Eigenvalue range & Count & Character \\
\midrule
Top positive & $+10$ to $+109$ & $1$ & grows with $\lambda$ \\
Near-zero & $-8 \times 10^{-3}$ to $-6 \times 10^{-8}$ & $67$--$94$ & the Lorentzian content \\
Bulk negative & $-0.04$ to $-61$ & $15$--$16$ & \textbf{constant count} \\
\bottomrule
\end{tabular}
\end{center}

On the near-zero subspace $V_0$ (dimension~$d_0$), $L$ and $D$ are \textbf{spectral mirrors}: $L|_{V_0}$ has signature $(d_0-1, 1)$ while $D|_{V_0}$ has signature $(1, d_0-1)$, and they cancel to extraordinary precision:
\begin{equation}
\frac{\|L|_{V_0} + D|_{V_0}\|}{\|L|_{V_0}\|} \sim 10^{-3}\text{--}10^{-5},
\end{equation}
with the cancellation \emph{improving} as $\lambda$ grows: from $0.24\%$ at $\lambda^2 = 50$ to $0.0036\%$ at $\lambda^2 = 10{,}000$.
\end{observation}

\begin{observation}[Prolate null space]\label{obs:prolate}
The near-zero subspace coincides with the prolate null space: $d_0 \approx 2N+1 - 2\pi N/L$, the dimensions outside the time-bandwidth product. The $\sim 15$ bulk eigenvalues occupy the prolate \emph{signal} space.
\end{observation}

\begin{observation}[Phase transition at $t = 1$]\label{obs:phase}
In the one-parameter family $M(t) = L + t\,D$, the number of positive eigenvalues undergoes a sharp phase transition:

\begin{center}
\begin{tabular}{rl}
\toprule
$t$ & \# positive eigenvalues \\
\midrule
$0.0$ & $69$ \\
$0.5$ & $68$ \\
$0.9$ & $68$ \\
$0.95$ & $67$ \\
$1.0$ & $\mathbf{1}$ \\
$1.01$ & $2$ \\
\bottomrule
\end{tabular}
\end{center}

Of $67$ eigenvalue zero-crossings, $66$ occur in $[0.95, 1.00)$---a single histogram spike. The transition point $t = 1.0000$ is \textbf{exact at all tested~$\lambda$} ($\lambda^2 = 50$ through $5{,}000$). The second eigenvalue has a parabolic minimum at $t = 1$ (positive on both sides), achieved through an eigenvector swap---the second eigenvector at $t = 0.99$ has overlap $0.006$ with the second eigenvector at $t = 1.01$.
\end{observation}

The spectral mirror has two consequences. First, the Lorentzian property is a \emph{collective} phenomenon: all $\sim 66$ near-zero eigenvalues scale linearly as $c_k \cdot (1 - t)$, crossing zero simultaneously at $t = 1$. Second, the effective rank of~$M$ at $1\%$ threshold is approximately~$15$, \emph{constant} as matrix dimension grows from $47$ to $111$---the prolate signal space carries all significant spectral content.

The mirror is $100\%$ prime-specific: Cram\'er-model primes produce zero near-zero eigenvalues ($31 \pm 5$ positive eigenvalues in every trial). Adding actual primes cumulatively, the near-zero cluster \emph{emerges gradually}: it appears at $\sim 10$ primes, grows linearly, and reaches full size only with all primes included.

%======================================================================
\section{The Displacement Identity}\label{sec:displacement}
%======================================================================

\begin{proposition}[Rank-$2$ displacement]\label{prop:displacement}
Let $D_n = \diag(-N, \ldots, N)$ be the position operator. Then
\begin{equation}\label{eq:displacement}
[D_n, M] := D_n M - M D_n = \mathbf{1}\, \mathbf{B}^T - \mathbf{B}\, \mathbf{1}^T,
\end{equation}
where $\mathbf{1} = (1, \ldots, 1)^T$ and $\mathbf{B} = (B_{-N}, \ldots, B_N)^T$ from~\eqref{eq:Bn}. The right side has rank exactly~$2$.
\end{proposition}

\begin{proof}
For $n \neq m$: $([D_n, M])_{nm} = n M_{nm} - M_{nm} m = (n-m) \cdot (B_m - B_n)/(n-m) = B_m - B_n = (\mathbf{1}\, \mathbf{B}^T - \mathbf{B}\, \mathbf{1}^T)_{nm}$. The diagonal is zero on both sides since $M$ is symmetric. Verified to relative precision $< 4 \times 10^{-15}$ at all tested parameters.
\end{proof}

The generators $\mathbf{1}$ (even) and $\mathbf{B}$ (odd) live in orthogonal parity subspaces, so the displacement couples only the even--odd cross terms. Within each parity block, $M$ commutes with the position operator projected to that block---a strong constraint that is trivially satisfied because $D_n$ maps even functions to odd and vice versa.

This is the Sylvester-type analog of the Lyapunov displacement $CD + DC = \mathbf{v}\mathbf{v}^T$ studied by Pushnitski--Treil \cite{PT2025} for weighted Cauchy matrices $C_{jk} = \sqrt{A_j A_k}/(a_j + a_k)$. Their spectral map is an involution; ours is not (the rank-$2$ antisymmetric structure maps to a different matrix class under spectral duality), but the displacement identity~\eqref{eq:displacement} is exact and encodes the Weil explicit formula in a single algebraic equation.

%======================================================================
\section{Conclusions}
%======================================================================

We have formulated the Lorentzian Weil Matrix Conjecture (Conjecture~\ref{conj:main}), proved it implies~RH (Theorem~\ref{thm:implies}), and developed a sign lemma (Theorem~\ref{thm:sign}) that detects off-line zeros with sensitivity $\delta_{\mathrm{crit}} \approx 1.4 \times 10^{-6}$---three orders of magnitude beyond the Vinogradov--Korobov zero-free region.

The kernel-vanishing property $K(L) = 0$ eliminates exponential growth of the zero-tail perturbation, yielding an $L$-independent bound (Remark~\ref{rem:KL0}). Combined with verified zeros up to height $3 \times 10^{12}$ and computational verification of $Q_W \geq 0$ to $\lambda^2 = 287{,}000$, the effective verification range extends to $\lambda^2 < e^{240{,}000}$ with a safety factor of $80{,}000$ (Theorem~\ref{thm:tail}).

The locking constant $\mathcal{Z} = \xi''(\tfrac{1}{2})/\xi(\tfrac{1}{2})$ encodes the total sensitivity of the Lorentzian property to zero displacement. It arises from an $80{:}1$ cancellation between the archimedean contribution ($A = -8 + \pi^2/4 + 2G$, involving only~$\pi$ and Catalan's constant) and the prime contribution ($P \approx 3.747$, involving~$\zeta$ at the critical point). The locking identity~\eqref{eq:lock} reformulates~RH as the equality $\mathcal{Z} = \sum 2/\gamma_k^2$---each zero contributes maximally when on the critical line. This identity is verifiable to arbitrary precision: the left side through the eta function, the right side through the computed zeros.

Extensive diagnostics establish that the Lorentzian property cannot be proved from the Cauchy--Loewner structure alone (the Cram\'er model gives $0/150$ negative-definite trials), nor from the geometric Hodge framework (which maps cleanly but hits the same quantitative wall), nor from any convergent prime sum (the natural boundary of the prime zeta function at $\mathrm{Re}(s) = 1$ prevents continuation to the critical point), nor from matrix-geometric tools. Specifically, the Bharali--Holtz Loewner inertia theorem (which gives Lorentzian signature for divided differences of~$x^r$, $1 < r < 2$) does not apply because $B_n \sim n^{-1}$; the Bezoutian root-counting framework requires displacement rank~$2$, but the odd block has displacement rank~$14$; and each prime power contributes a full-rank perturbation to~$M$, precluding eigenvalue tracking via Cauchy interlacing. The Lorentzian property encodes the specific distribution of primes enforced by the zeros of~$\zeta$, and any proof must engage with this arithmetic content.

The decomposition of the Tur\'an ratios (Section~\ref{sec:decomp_turan}) offers a sharper path. The archimedean prefactor~$F$ satisfies the Tur\'an inequalities on its own, with margin decaying geometrically as $\sim 3^{-k}$. Multiplication by~$\zeta$ does not destroy this margin but \emph{amplifies} it, with the boost growing as~$k$ increases. This reduces~RH to two concrete steps: (1)~prove $R_k^{(F)} > 1$ for all~$k$ from Gamma-function asymptotics (a purely archimedean statement), and (2)~prove that the zeta correction preserves log-concavity (the arithmetic content). Step~1 requires no number theory; Step~2 encodes the essential interplay between primes and the functional equation.

The spectral mirror (Section~\ref{sec:mirror}) provides the sharpest structural explanation of the Lorentzian property: on the prolate null space ($\sim 80\%$ of dimensions), the Cauchy off-diagonal and archimedean diagonal are spectral conjugates cancelling to $< 0.1\%$, with the cancellation improving as $\lambda$ grows. The phase transition at $t = 1.0$ (Observation~\ref{obs:phase})---66 eigenvalues crossing zero simultaneously in a collective avalanche at the exact physical parameter---reveals the Lorentzian property as a critical phenomenon. However, the prolate null space cancellation is the \emph{easy} part: the hard content of~RH lives in the complementary subspace where $L$ and $D$ \emph{reinforce} rather than cancel. The dimension of this ``signal space'' grows with the truncation parameter~$N$ (scaling as $\sim 0.12 \, N^{1.33}$), but the Lorentzian signature and negative definiteness of~$M_{\mathrm{odd}}$ are $N$-converged structural invariants, independent of truncation.

The exact displacement identity $[D_n, M] = \mathbf{1}\mathbf{B}^T - \mathbf{B}\mathbf{1}^T$ (Proposition~\ref{prop:displacement}) encodes the Weil explicit formula as a rank-$2$ commutator equation, connecting the Fourier-index operator to the Cauchy generating function~$B_n$. This is the Sylvester-type analog of the Lyapunov displacement for Pushnitski--Treil weighted Cauchy matrices \cite{PT2025}, but with orthogonal parity generators that prevent the spectral involution from closing.

The structural parallel with misaligned supersymmetry (Remark~\ref{rem:susy}) illuminates both steps: modular invariance of partition functions forces Bose--Fermi cancellations to leave non-negative residuals in non-supersymmetric gauge theories \cite{Dienes1994}. The functional equation of~$\zeta$ is the number-theoretic analogue of modular invariance. In both settings, a symmetry constraint forces near-cancellation with a positive residual, and the positivity of the residual has deep structural consequences.

%======================================================================
\section*{Acknowledgments}
%======================================================================

This work was conducted over 77 computational sessions using the Connes--Consani--Moscovici framework, with literature review assisted by automated research agents. All numerical results are reproducible from the Python scripts in the project repository.

%======================================================================
\begin{thebibliography}{99}
%======================================================================

\bibitem{AHK2018}
K.~Adiprasito, J.~Huh, and E.~Katz,
\emph{Hodge theory for combinatorial geometries},
Ann.\ of Math.\ \textbf{188} (2018), 381--452.
arXiv:1511.02888.

\bibitem{BH2020}
P.~Br\"and\'en and J.~Huh,
\emph{Lorentzian polynomials},
Ann.\ of Math.\ \textbf{192} (2020), 821--891.

\bibitem{CC2020}
A.~Connes and C.~Consani,
\emph{Weil positivity and trace formula, the archimedean place},
Selecta Math.\ (2021).
arXiv:2006.13771.

\bibitem{CC2025}
A.~Connes and C.~Consani,
\emph{Zeta spectral triples},
preprint (2025).
arXiv:2511.22755.

\bibitem{CCM2025}
A.~Connes,
\emph{The Riemann Hypothesis: past, present, and a letter through time},
preprint (2026).
arXiv:2602.04022.

\bibitem{Markovsky2014}
I.~Markovsky,
\emph{Sum-of-exponentials via Hankel low-rank approximation},
preprint (2014).

\bibitem{Morishita2025}
M.~Morishita,
\emph{On a relation between Deninger's foliated dynamical systems and Connes--Consani's adelic spaces},
M\"unster J.\ Math.\ (2025).
arXiv:2508.15971.

\bibitem{Basar2019}
G.~Ba\c{s}ar, A.~Cherman, and D.~McGady,
\emph{Bose--Fermi cancellations without supersymmetry},
Phys.\ Rev.\ D \textbf{99} (2019), 065014.
arXiv:1812.04642.

\bibitem{Dienes1994}
K.~R.~Dienes,
\emph{Modular invariance, finiteness, and misaligned supersymmetry},
Nucl.\ Phys.\ B \textbf{429} (1994), 533--588.
arXiv:hep-th/9402006.

\bibitem{GORZ2019}
M.~Griffin, K.~Ono, L.~Rolen, and D.~Zagier,
\emph{Jensen polynomials for the Riemann zeta function and other sequences},
Proc.\ Natl.\ Acad.\ Sci.\ USA \textbf{116} (2019), 11103--11110.

\bibitem{ConnesChamseddine1996}
A.~Connes and A.~Chamseddine,
\emph{The spectral action principle},
Comm.\ Math.\ Phys.\ \textbf{186} (1997), 731--750.
arXiv:hep-th/9606001.

\bibitem{Platt2021}
D.~Platt and T.~Trudgian,
\emph{The Riemann hypothesis is true up to $3 \times 10^{12}$},
Bull.\ London Math.\ Soc.\ \textbf{53} (2021), 792--797.

\bibitem{Voros2003}
A.~Voros,
\emph{Zeta functions for the Riemann zeros},
Ann.\ Inst.\ Fourier \textbf{53} (2003), 665--699.

\bibitem{PT2025}
A.~Pushnitski and S.~Treil,
\emph{The spectral map for weighted Cauchy matrices is an involution},
preprint (2025).
arXiv:2504.18707.

\bibitem{BharaliHoltz2015}
G.~Bharali and O.~Holtz,
\emph{Functions preserving nonnegativity of matrices},
SIAM J.\ Matrix Anal.\ Appl.\ \textbf{30} (2009), 13--32.
arXiv:1501.01505.

\bibitem{Bombieri2000}
E.~Bombieri,
\emph{Remarks on Weil's quadratic functional in the theory of prime numbers},
Atti Accad.\ Naz.\ Lincei Cl.\ Sci.\ Fis.\ Mat.\ Nat.\ Rend.\ Lincei \textbf{11} (2000), 183--233.

\bibitem{CNV1986}
G.~Csordas, T.~S.~Norfolk, and R.~S.~Varga,
\emph{The Riemann hypothesis and the Tur\'an inequalities},
Trans.\ Amer.\ Math.\ Soc.\ \textbf{296} (1986), 521--541.

\bibitem{RoyKailath1989}
R.~Roy and T.~Kailath,
\emph{ESPRIT---estimation of signal parameters via rotational invariance techniques},
IEEE Trans.\ ASSP \textbf{37} (1989), 984--995.

\end{thebibliography}

\end{document}
