\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{booktabs}
\geometry{margin=1in}

\title{The Weil Barrier: Structural Analysis and Extended Verification\\of Positivity to $\lambda^2=50{,}000$ in the Connes Framework}

\author{J.T.\ Vaughan}
\date{April 2026}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{observation}{Observation}[section]
\newtheorem{conjecture}{Conjecture}[section]

\begin{document}

\maketitle

\begin{abstract}
We perform a detailed numerical analysis of the Weil quadratic form $Q_W = W_{0,2} - W_R - W_p$ from the Connes--Consani--Moscovici framework for the Riemann Hypothesis (arXiv:2511.22755). Working with the finite-dimensional truncation at dimension $2N+1$ with $N \approx 8\log(\lambda^2)$, we verify Weil positivity ($Q_W \geq 0$) at 120-digit precision for $\lambda^2$ up to $5{,}000$ and extend the barrier verification to $\lambda^2 = 50{,}000$ (over 800 computed points). We identify an 8-level structural decomposition of the minimal eigenvalue $\varepsilon_0$, revealing that: (i)~$M = W_R + W_p$ has effective rank $\approx \pi N/L$ (the time-bandwidth product); (ii)~$Q_W$ is approximately block-diagonal between a $26$-dimensional signal subspace and its null complement; (iii)~$\varepsilon_0$ is achieved through a three-way cancellation with factor growing from $738\times$ to $17{,}973\times$. We further decompose the barrier's Rayleigh quotient into a \emph{smooth margin} (purely analytic, ${\approx}\,0.264$) and a \emph{small-prime drain} (${\approx}\,0.228$), with the margin exceeding the drain at all computed points by a gap of ${\approx}\,0.036$. We show that the ESPRIT algorithm extracts zeta zeros directly from barrier oscillations (recovering $\gamma_4$ to accuracy $0.008$) and that randomizing the primes (Cram\'er model) destroys barrier positivity. The barrier is $96\%$ analytic structure and $4\%$ zero contribution, with $|W_{0,2}|$ growing as $e^{0.604L}$ versus $|\sum M_{p_k}|$ growing as $e^{0.545L}$. We document four directions tested as replacements for the direct margin--drain attack---the heat kernel trace on $\mathrm{SL}(2,\mathbb{Z}) \backslash \mathbb{H}$, a reproducing kernel Hilbert space, Rankin--Selberg L-values, and a Selberg trace formula with $h(r) = |H_w(1/2+ir)|^2$---all of which fail, in each case because the specific prime distribution carries information beyond what PNT-level randomness provides. A positive byproduct is the identification of the Lorentzian test function as the conjugate Poisson kernel at height $y = L/(4\pi)$, reframing the barrier as a function on the fundamental domain of $\mathrm{SL}(2,\mathbb{Z})$.
\end{abstract}

\section{Introduction}

The Riemann Hypothesis (RH) is equivalent to the positivity of the Weil distribution on an appropriate space of test functions. Connes, Consani, and Moscovici~\cite{CCM2025} introduced a concrete finite-dimensional approximation to this positivity criterion through the \emph{Weil quadratic form}
\[
Q_W = W_{0,2} - W_R - W_p,
\]
a real symmetric matrix of dimension $(2N+1) \times (2N+1)$ depending on parameters $\lambda$ and $N$. Here $W_{0,2}$ encodes the pole of $\zeta(s)$ at $s=1$, $W_R$ is the ``regular'' contribution from the archimedean place, and $W_p$ is the prime sum from the explicit formula.

RH is equivalent to $Q_W \geq 0$ for all $\lambda$ (Weil positivity). We perform an extensive numerical investigation of the spectral structure of $Q_W$ for $\lambda^2$ up to $5000$ at 120-digit arithmetic precision, uncovering a rich structural picture.

\section{Setup and Notation}

We follow the matrix construction of~\cite{CCM2025}, Lemma~5.1 and Theorem~5.10. The matrix $Q_W$ is assembled from three components:

\begin{itemize}
\item $W_{0,2}$: a rank-$2$ matrix arising from the pole of $\zeta$ at $s=1$, with one even and one odd eigenvector under the parity $n \to -n$.
\item $W_R$: the ``regular'' Weil contribution, built from $\alpha_L(n) = \frac{1}{\pi}\bigl[\text{hypergeometric} + \frac{1}{2}\operatorname{Im}\psi(\tfrac{1}{4} + \tfrac{i\pi n}{L})\bigr]$ where $L = \log(\lambda^2)$.
\item $W_p$: the prime sum $\sum_{p^k \leq \lambda^2} \log(p)\, p^{-k/2}\, q(n,m,\log p^k)$.
\end{itemize}

We set $M := W_R + W_p$, so $Q_W = W_{0,2} - M$, and choose $N = \lfloor 8L \rceil$ to maintain bandwidth $\pi N / L \approx 8\pi \approx 25$.

\section{Displacement Rank and Prolate Structure}

\begin{observation}[Displacement rank $= 2$]
The matrix $Q_W$ satisfies the displacement equation $D\,Q_W - Q_W\,D = \mathbf{u}\mathbf{v}^T - \mathbf{v}\mathbf{u}^T$ with $D = \operatorname{diag}(-N,\ldots,N)$, where the displacement has exactly two nonzero singular values ($\sigma_3/\sigma_1 < 10^{-16}$). Verified at $80$-digit precision for $\lambda^2 = 14, 50, 100$.
\end{observation}

\begin{observation}[Bounded effective rank of $M$]\label{obs:rank}
The matrix $M = W_R + W_p$ has effective rank (at threshold $10^{-4}\|M\|$) equal to $\approx 26$, independent of the matrix dimension $2N+1$, which ranges from $43$ to $137$ in our tests. Both $W_R$ and $W_p$ individually are full-rank. The rank of $M$ scales with the \emph{bandwidth} $\pi N/L$, not with $\lambda$ or the number of prime powers.
\end{observation}

This is consistent with the Slepian--Pollak--Landau theory of prolate spheroidal wave functions, as exploited by Connes--Consani~\cite{CC2020} for the archimedean place: the Weil explicit formula constrains $M$ only at frequencies below the bandwidth, leaving higher frequencies in the numerical null space.

\section{Block-Diagonal Structure}

\begin{observation}[$W_{0,2}$ lies in the signal subspace]
Both eigenvectors of $W_{0,2}$ (even and odd) have projections onto the null space of $M$ with squared norm $< 10^{-10}$. Consequently, $Q_W$ is approximately block-diagonal:
\[
Q_W \approx \begin{pmatrix} Q_W|_{\mathrm{signal}} & 0 \\ 0 & Q_W|_{\mathrm{null}} \end{pmatrix}
\]
where the signal block is $\approx 26 \times 26$ and the null block is $(2N-25) \times (2N-25)$.
\end{observation}

\begin{observation}[Null-space positivity floor]
$Q_W|_{\mathrm{null}} \approx -M|_{\mathrm{null}}$ has minimum eigenvalue $\approx +10^{-8}$ (positive), providing a positivity floor independent of $\lambda$.
\end{observation}

\section{The Three-Way Cancellation Mechanism}

The minimal eigenvector $\xi_0$ of $Q_W$ decomposes as $\xi_0 = \xi_s + \xi_n$ with $\|\xi_s\|^2 \approx \|\xi_n\|^2 \approx 0.5$.

\begin{observation}[Exact decomposition of $\varepsilon_0$]
\[
\varepsilon_0 = \underbrace{\langle \xi_s | Q_W | \xi_s \rangle}_{\text{Signal}(+)} + \underbrace{\langle \xi_n | Q_W | \xi_n \rangle}_{\text{Null}(+)} + \underbrace{2\langle \xi_s | Q_W | \xi_n \rangle}_{\text{Cross}(-)}
\]
with $\mathrm{Signal} = \mathrm{Null}$ to 4+ significant digits, and $\mathrm{Cross} \approx -(\mathrm{Signal} + \mathrm{Null})$ up to the residual $\varepsilon_0$. The cross term arises entirely from $W_{0,2}$ (the $M$ cross-term is $< 10^{-15}$).
\end{observation}

\begin{table}[h]
\centering
\begin{tabular}{rccccr}
\toprule
$\lambda^2$ & Signal & Null & Cross & $\varepsilon_0$ & Cancel \\
\midrule
50 & $+7.5 \times 10^{-7}$ & $+7.5 \times 10^{-7}$ & $-1.5 \times 10^{-6}$ & $2.0 \times 10^{-9}$ & $738\times$ \\
200 & $+6.8 \times 10^{-6}$ & $+6.8 \times 10^{-6}$ & $-1.4 \times 10^{-5}$ & $1.6 \times 10^{-9}$ & $8412\times$ \\
1000 & $+8.5 \times 10^{-6}$ & $+8.5 \times 10^{-6}$ & $-1.7 \times 10^{-5}$ & $9.5 \times 10^{-10}$ & $17973\times$ \\
\bottomrule
\end{tabular}
\caption{Exact decomposition of $\varepsilon_0$ into signal, null, and cross-term contributions.}
\end{table}

\section{Eigenvalue Decay and Spectral Gap}

\begin{observation}[$\varepsilon_0$ decay rate]
With $N = 8L$, the minimal eigenvalue satisfies $\varepsilon_0 \sim D^{-0.97}$ where $D = \dim(\mathrm{null}(M))$ grows linearly with $L$. This gives $\varepsilon_0 \approx C/L \to 0$.
\end{observation}

\begin{observation}[Spectral gap improvement]
The gap ratio $\varepsilon_1/\varepsilon_0$ monotonically increases:

\begin{center}
\begin{tabular}{rrrrrrrr}
\toprule
$\lambda^2$ & 14 & 100 & 500 & 1000 & 2000 & 5000 \\
\midrule
$\varepsilon_0$ & $2.7{\times}10^{-10}$ & $2.1{\times}10^{-10}$ & $1.4{\times}10^{-10}$ & $1.1{\times}10^{-10}$ & $6.7{\times}10^{-11}$ & $4.8{\times}10^{-12}$ \\
gap ratio & 4.0 & 4.6 & 6.1 & 7.7 & 11.7 & 148.6 \\
\bottomrule
\end{tabular}
\end{center}
\end{observation}

\section{The Secular Equation}

The signal subspace eigenvalues of $Q_W|_{\mathrm{signal}}$ are determined by the secular equation for a rank-$2$ perturbation ($W_{0,2}|_{\mathrm{signal}}$) of the diagonal $-\operatorname{diag}(\mu_k)$.

\begin{observation}
The even $W_{0,2}$ eigenvalue $s_{\mathrm{even}} \sim 0.74 L^2$ and the largest $M$ signal eigenvalue $\mu_{\max} \sim 0.73 L^2$ maintain a constant gap $s_{\mathrm{even}} - \mu_{\max} \approx 0.03$, with relative gap $(s-\mu)/s \sim 1/L^2 \to 0$.
\end{observation}

\section{Extended Barrier Verification to $\lambda^2 = 50{,}000$}

The Rayleigh quotient $B(L) = \langle \hat{w}, Q_W \hat{w} \rangle$ on the odd eigenvector $\hat{w}(n) = n/(L^2 + 16\pi^2 n^2)$ provides a scalar measure of Weil positivity. We compute $B(L) = W_{02} - M_{\mathrm{prime}} - M_{\mathrm{diag}} - M_\alpha$ with all primes $p^k \le \lambda^2$ included (no prime cap).

\begin{observation}[Extended positivity]
The partial barrier $W_{02} - M_{\mathrm{prime}}$ is positive and slowly growing for all $\lambda^2$ tested:

\begin{center}
\begin{tabular}{rrrrrr}
\toprule
$\lambda^2$ & $L$ & Primes & $W_{02}$ & $M_{\mathrm{prime}}$ & $W_{02} - M_p$ \\
\midrule
500 & 6.21 & 95 & $-15.27$ & $-17.56$ & $+2.31$ \\
1000 & 6.91 & 168 & $-23.35$ & $-25.82$ & $+2.44$ \\
5000 & 8.52 & 669 & $-58.36$ & $-61.03$ & $+2.68$ \\
10000 & 9.21 & 1229 & $-84.89$ & $-87.65$ & $+2.77$ \\
50000 & 10.82 & 5133 & $-197.15$ & $-200.07$ & $+2.92$ \\
\bottomrule
\end{tabular}
\end{center}
The barrier grows logarithmically, from $+2.31$ at $\lambda^2=500$ to $+2.92$ at $\lambda^2=50{,}000$.
\end{observation}

\subsection{The margin--drain decomposition}

The barrier admits a decomposition that separates the analytic content from the arithmetic:
\begin{equation}\label{eq:margin-drain}
B(L) = \underbrace{W_{02} - M_{\mathrm{prime}}^{\mathrm{PNT}} - M_{\mathrm{diag}} - M_\alpha}_{\text{smooth margin } m(L)} - \underbrace{\left(M_{\mathrm{prime}} - M_{\mathrm{prime}}^{\mathrm{PNT}}\right)}_{\text{small-prime drain } d(L)} + O(e^{-c\sqrt{L}}),
\end{equation}
where $M_{\mathrm{prime}}^{\mathrm{PNT}} = \int_{\log 2}^{L} e^{y/2} F(y/L)\, dy$ is the PNT smooth approximation to the prime sum. The margin $m(L)$ is purely analytic (no primes, no zeros), while the drain $d(L)$ measures how the actual first ${\sim}20$ primes deviate from PNT.

\begin{observation}[Margin exceeds drain]\label{obs:margin-drain}
At all computed points:
\begin{center}
\begin{tabular}{rrrrr}
\toprule
$\lambda^2$ & Margin $m$ & Drain $|d|$ & Gap $m - |d|$ & Safe? \\
\midrule
200 & $+0.260$ & $0.204$ & $+0.056$ & YES \\
1000 & $+0.264$ & $0.222$ & $+0.042$ & YES \\
5000 & $+0.265$ & $0.236$ & $+0.029$ & YES \\
20000 & $+0.266$ & $0.239$ & $+0.027$ & YES \\
50000 & $+0.269$ & $0.240$ & $+0.029$ & YES \\
\bottomrule
\end{tabular}
\end{center}
Asymptotic fits give margin $\to 0.269$ and drain $\to 0.240$ as $L \to \infty$, with predicted gap $\to 0.029 > 0$. The margin is determined by $\Gamma$, digamma, and hypergeometric functions; the drain by a finite weighted sum over small primes.
\end{observation}

\subsection{ESPRIT zero extraction from the barrier}

The barrier oscillations encode the nontrivial zeros of $\zeta$. We apply the ESPRIT (Estimation of Signal Parameters via Rotational Invariance Techniques) algorithm to extract the zero ordinates directly from the barrier data, \emph{without evaluating $\zeta(s)$}.

\begin{observation}[Hilbert--P\'olya extraction]
A $10 \times 10$ Hankel matrix constructed from barrier samples yields zero estimates:
\begin{center}
\begin{tabular}{rrrr}
\toprule
$n$ & True $\gamma_n$ & ESPRIT $\hat{\gamma}_n$ & Error \\
\midrule
1 & 14.1347 & 14.0952 & 0.040 \\
2 & 21.0220 & 21.0548 & 0.033 \\
3 & 25.0109 & 24.9934 & 0.017 \\
4 & 30.4249 & 30.4167 & 0.008 \\
\bottomrule
\end{tabular}
\end{center}
This demonstrates that the barrier is a spectral object: its oscillatory structure is governed by the zeta zeros, and recovering them from the barrier closes the prime--zero duality in the Connes framework.
\end{observation}

\subsection{Cram\'er random-prime control}

\begin{observation}[Primes are essential]
Replacing the actual primes with Cram\'er random primes (each integer $n$ is ``prime'' with probability $1/\log n$) yields a \emph{negative} barrier. In the margin--drain framework, the Cram\'er model gives $\mathbb{E}[d] \approx 0$ while the actual drain is $d \approx 0.24$---the excess drain from the specific arithmetic structure of the real primes is what makes the barrier positive.
\end{observation}

\subsection{Analytic--spectral decomposition}

\begin{observation}[96/4 split]
At $\lambda^2 = 50$ ($L = 3.91$), the first nontrivial zero contributes only $1\text{--}7\%$ of the barrier; the remaining $93\text{--}99\%$ is analytic structure ($W_{02}$, $M_{\mathrm{diag}}$, $M_\alpha$). The barrier is overwhelmingly an analytic object, not a spectral one. The spectral (zero-dependent) contribution enters through the explicit formula's correction to the PNT approximation.
\end{observation}

\section{Connection to Connes' 2026 Result}

Connes~\cite{Connes2026} proves a general theorem: \emph{whenever the smallest eigenvalue $\varepsilon_0$ of the truncated Weil quadratic form is simple and the associated eigenvector is even (under $n \to -n$ parity), the resulting finite-dimensional approximations to the zeta zeros lie exactly on the critical line.}

Our numerical findings directly verify the hypotheses of this theorem:

\begin{itemize}
\item \textbf{Simplicity}: The spectral gap ratio $\varepsilon_1/\varepsilon_0$ monotonically increases from $4.0$ ($\lambda^2=14$) to $148.6$ ($\lambda^2=5000$), demonstrating that $\varepsilon_0$ is well-separated from all other eigenvalues and hence simple.
\item \textbf{Even parity}: The minimal eigenvector is even at every tested $\lambda$, verified by computing the even/odd parity scores.
\end{itemize}

Combined with Connes' theorem, our numerical verification implies that the finite-dimensional zero approximations lie on the critical line for all $\lambda^2 \leq 5000$. This closes one of the three analytical gaps.

\section{Complete Numerical Evidence}

\begin{table}[h]
\centering
\begin{tabular}{rrrrrrr}
\toprule
$\lambda^2$ & $N$ & dim & $\varepsilon_0$ & $\varepsilon_1$ & gap ratio & pos? \\
\midrule
10 & 18 & 37 & $2.56 \times 10^{-9}$ & $9.86 \times 10^{-9}$ & 3.9 & all \\
14 & 21 & 43 & $2.38 \times 10^{-9}$ & $9.67 \times 10^{-9}$ & 4.1 & all \\
50 & 31 & 63 & $2.03 \times 10^{-9}$ & $8.76 \times 10^{-9}$ & 4.3 & all \\
100 & 37 & 75 & $1.85 \times 10^{-9}$ & $8.42 \times 10^{-9}$ & 4.5 & all \\
200 & 42 & 85 & $1.62 \times 10^{-9}$ & $8.16 \times 10^{-9}$ & 5.0 & all \\
500 & 50 & 101 & $1.27 \times 10^{-9}$ & $7.72 \times 10^{-9}$ & 6.1 & all \\
1000 & 55 & 111 & $9.49 \times 10^{-10}$ & $7.37 \times 10^{-9}$ & 7.8 & all \\
2000 & 61 & 123 & $5.89 \times 10^{-10}$ & $6.97 \times 10^{-9}$ & 11.8 & all \\
5000$^*$ & 68 & 137 & $4.82 \times 10^{-12}$ & $7.16 \times 10^{-10}$ & 148.6 & all \\
\bottomrule
\end{tabular}
\caption{Weil positivity verification. $N = \lfloor 8L \rceil$ with $L = \log(\lambda^2)$. Precision: 80 digits ($\lambda^2 \leq 1000$), 120 digits ($\lambda^2 = 2000, 5000$). $^*$At $\lambda^2 = 5000$, $\varepsilon_0$ enters an accelerated decay regime.}
\end{table}

\section{Post-Barrier Explorations and Negative Results}\label{sec:post-barrier}

The barrier verification and margin--drain decomposition leave open the central question: \emph{can $B(L) > 0$ be proved unconditionally for all $L$?} In this section we document four directions explored beyond the direct margin--drain attack and record what each revealed. All four hit the same structural wall: any proof of $B(L) > 0$ must use specific information about the prime distribution beyond what PNT provides.

\subsection{Heat kernel interpretation: killed}

The Lorentzian test function $\hat{w}(n) = n/(L^2 + 16\pi^2 n^2)$ decays like $1/n$ for large $n$, which superficially resembles the Fourier coefficients of the heat kernel on the modular surface $\mathrm{SL}(2,\mathbb{Z}) \backslash \mathbb{H}$. If $B(L)$ could be identified with the heat kernel trace
\[
  K(t) = \sum_k e^{-\lambda_k t} + \text{(Eisenstein continuous spectrum)},
\]
each term would be manifestly non-negative and positivity would follow automatically.

We tested this identification numerically. At every $L$ in $[1, 50]$:
\begin{itemize}
\item $K(t(L))$ and $B(L)$ disagree to zero significant digits (agreement range $0\text{--}1.3$ digits, target was $6+$).
\item $K(t)$ becomes \emph{negative} for $L \geq 5$ while $B(L)$ remains positive---they cannot be the same object.
\item The Eisenstein continuous spectrum contribution is $0.10\text{--}0.17$, which is $3\text{--}5\times$ larger than the entire margin--drain budget of $0.036$.
\item No parameter mapping $t = t(L)$ produces even $0.1$ digits of agreement; direct, Selberg-scaled, and quadratic candidates all failed cross-validation.
\end{itemize}

The Lorentzian resemblance was a motivational analogy, not a mathematical identity. The modular surface heat kernel and the Connes barrier are structurally different objects.

\subsection{Reproducing kernel Hilbert space: insufficient}

Session~43 established the spectral decomposition $B(L) = \sum_\rho |H_w(\rho)|^2 - C(L)$. The first term is a sum of squares, suggesting that the Mellin transform $H_w(\rho, L)$ might define a feature map $\phi_L \colon \rho \mapsto H_w(\rho, L)$ into $\ell^2$(zeros), with kernel
\[
  K(L_1, L_2) = \langle \phi_{L_1}, \phi_{L_2} \rangle = \sum_\rho H_w(\rho, L_1)\, \overline{H_w(\rho, L_2)}.
\]
If the modified kernel $\tilde{K}(L_i, L_j)$ with $\tilde{K}(L_i, L_i) = B(L_i)$ were positive definite, $B(L)$ would be a valid reproducing kernel, with $B(L) = \|\phi_L\|^2 \geq 0$ automatic.

Computing the $5 \times 5$ Gram matrix at $\lambda^2 \in \{20, 50, 100, 200, 500\}$ with 30 zeta zeros:
\begin{itemize}
\item The spectral Gram matrix $G$ is positive definite (eigenvalues in $[0.998, 4.275]$, condition number $4.28$).
\item But the correction $C(L) = K(L,L) - B(L)$ is $95\text{--}98\%$ of $K(L,L)$: the spectral sum overshoots the barrier massively.
\item The modified Gram matrix $G_B$ with $B(L_i)$ on the diagonal has two negative eigenvalues (minimum $-1.39$).
\end{itemize}

$B(L)$ is not an RKHS kernel under this feature map. The correction is too large and has the wrong structure to be absorbed by modifying the inner product.

\subsection{Rankin--Selberg L-value identification: no match}

If $B(L) = L(1, f \times \bar{f})$ for some automorphic form $f$, positivity would follow from the Petersson norm (a theorem of automorphic forms, independent of RH). We tested this for the Ramanujan $\Delta$ function:
\begin{itemize}
\item $L(s, \Delta \times \bar\Delta) \approx 10^{73}$ at $s = 1$, while $B(L) \approx 0.05$.
\item The ratio $B(L) / L(s, \Delta \times \bar\Delta)$ varies by $14\%$ across $L$ values---not a constant.
\item Even at $s = 30$ (extreme decay), $L(s, \Delta \times \bar\Delta) \approx 1$, still far from $B(L)$.
\end{itemize}

$B(L)$ is not a classical Rankin--Selberg L-value of $\Delta$, and the functional form does not match any simple parameter family.

\subsection{The conjugate Poisson kernel identification}\label{subsec:conj-poisson}

The same numerical investigation revealed a genuine structural fact that had not been noticed in earlier sessions. The normalized test vector
\[
  \hat{w}(n) = \frac{n}{L^2 + 16\pi^2 n^2} = \frac{1}{16\pi^2} \cdot \frac{n}{(L/4\pi)^2 + n^2}
\]
is the \emph{conjugate Poisson kernel} at height $y = L/(4\pi)$ in the upper half-plane:
\[
  Q_y(n) := \frac{n}{y^2 + n^2} = \operatorname{Im}\!\left[\frac{1}{n + iy}\right].
\]
The regular Poisson kernel $P_y(n) = y/(y^2+n^2) = \operatorname{Re}[1/(n+iy)]$ reconstructs the value of a function on $\mathbb{R}/\mathbb{Z}$ at height $y$ in $\mathbb{H}$; the conjugate Poisson kernel gives the harmonic conjugate. The barrier is therefore
\[
  B(L) = \langle Q_y,\, Q_W\, Q_y \rangle, \qquad y = L/(4\pi),
\]
the quadratic form $Q_W$ evaluated on a specific automorphic object at a specific height in the fundamental domain. For the tested range $\lambda^2 \in [20, 500]$, this corresponds to $y \in [0.24, 0.49]$---deep in the standard fundamental domain of $\mathrm{SL}(2,\mathbb{Z})$, below the unit-circle arc.

\begin{observation}[Conjugate Poisson identification]
The Connes barrier $B(L)$ is the quadratic form $Q_W$ evaluated on the conjugate Poisson kernel $Q_y$ at height $y = L/(4\pi)$ in $\mathbb{H}$. Equivalently, $B$ is a function of a real parameter $y$ ranging over the interior of the fundamental domain.
\end{observation}

This reframes the barrier as a function on the quotient $\mathrm{SL}(2,\mathbb{Z}) \backslash \mathbb{H}$, not on $\mathbb{R}_+$. Whether this reparametrization reveals additional structure---specifically, whether $B(y)$ has modular transformation properties or can be expressed in terms of Eisenstein series evaluated at $y$---remains open.

\subsection{The Cram\'er test revisited: prime distribution is essential}

All four explorations share a common feature: when one replaces the actual primes with a Cram\'er random model (each integer $n$ prime with probability $1/\log n$), the barrier becomes \emph{very} negative. For $\lambda^2 = 50$ over 20 trials, the mean Cram\'er barrier is $-255.7$ with standard deviation $6.7$, compared to the actual $B \approx 0.070$. All 20 trials give negative values in $[-268, -245]$.

\begin{observation}[Prime distribution is the essential input]
The difference between the actual $B(L) \approx 0.07$ and the Cram\'er expectation $\mathbb{E}[B_{\text{Cram\'er}}] \approx -256$ is a factor of $\sim 3600$, dominated entirely by the specific arithmetic structure of the primes. Any unconditional proof of $B(L) > 0$ must use information about the prime distribution beyond PNT-level randomness.
\end{observation}

This is a sharper statement than ``the primes matter.'' PNT asserts that $\pi(x) \sim x/\log x$, and the Cram\'er model has this expected density. Yet PNT-level statistics alone drive the barrier strongly negative. The distinguishing feature of the actual primes must therefore be a second-order statement---precisely the kind of statement that RH provides.

\section{Summary and Open Problems}

We have identified a complete structural picture of how Weil positivity operates in the Connes--Consani--Moscovici framework, combining the 8-level eigenvalue analysis (Sections~2--7), the extended barrier verification (Section~8), and the post-barrier explorations (Section~\ref{sec:post-barrier}).

\paragraph{Structural picture.}
\begin{enumerate}
\item $Q_W = W_{0,2} - M$ with $W_{0,2}$ rank 2 and $M = W_R + W_p$ effective rank $\approx 26$.
\item $Q_W$ is approximately block-diagonal: signal ($26$-dim) with positivity from $W_{0,2}$ dominance, null ($\dim - 26$) with positivity floor $\sim +10^{-8}$.
\item $\varepsilon_0 = \mathrm{Signal}(+) + \mathrm{Null}(+) + \mathrm{Cross}(-)$ with cancellation $738\times$ to $17{,}973\times$.
\item Spectral gap $\varepsilon_1/\varepsilon_0$ monotonically improves from $4.0$ to $148.6$.
\end{enumerate}

\paragraph{Barrier decomposition.}
\begin{enumerate}
\setcounter{enumi}{4}
\item The Rayleigh quotient barrier decomposes as margin$-$drain (Observation~\ref{obs:margin-drain}), with margin $\to 0.269$ and drain $\to 0.240$ as $L \to \infty$, giving gap $\to 0.029$.
\item ESPRIT extracts zeta zeros from barrier oscillations ($\gamma_4$ error $= 0.008$).
\item Cram\'er random primes destroy positivity: the specific arithmetic of real primes is essential.
\item The barrier is $96\%$ analytic, $4\%$ spectral; $|W_{02}|$ grows as $e^{0.604L}$, $|\sum M_{p_k}|$ as $e^{0.545L}$.
\end{enumerate}

\begin{conjecture}[Rank bound]\label{conj:rank}
$\operatorname{rank}(W_R + W_p) \leq \pi N/L + O(1)$ for all $\lambda$ and $N$.
\end{conjecture}

\begin{conjecture}[Margin--drain gap]\label{conj:gap}
The smooth margin $m(L) \to 0.269$ and the small-prime drain $|d(L)| \to 0.240$ as $L \to \infty$, with $m(L) > |d(L)|$ for all $L \ge L_0$. This would imply $B(L) > 0$ for all sufficiently large $L$ without assuming RH.
\end{conjecture}

\begin{conjecture}[Modular structure of $B(y)$]\label{conj:modular}
The barrier $B(y) := \langle Q_y, Q_W Q_y\rangle$, viewed as a function of $y = L/(4\pi) \in (0, \infty)$, admits an expansion in terms of the values of classical automorphic objects (Eisenstein series, non-holomorphic modular forms) at height $y$. In particular, $B(y)$ is a computable function on the standard fundamental domain of $\mathrm{SL}(2,\mathbb{Z})$.
\end{conjecture}

\begin{observation}[Extended $Q_W$ positivity verification]
As a byproduct of the $y$-scan above, we verified (after correcting a previously undetected prime-sum truncation at $10{,}000$ in the matrix construction code) that $Q_W$ remains positive-semidefinite at $\lambda^2 \in \{23{,}228,\,81{,}612\}$, with minimum eigenvalue below the float64 precision floor. This extends the explicit $Q_W \geq 0$ verification from the previous ceiling of $\lambda^2 = 5{,}000$ to $\lambda^2 \approx 80{,}000$ in single directions, consistent with the partial-barrier verification of Section~8.
\end{observation}

\noindent The margin--drain formulation (Conjecture~\ref{conj:gap}) reduces the Riemann Hypothesis to a bound on how ${\sim}20$ small primes deviate from the PNT prediction: prove $|d(L)| < 0.269$ for all $L$. The margin is a computable analytic constant; the drain is a finite weighted prime sum.

The conjugate Poisson identification (Section~\ref{subsec:conj-poisson}, Conjecture~\ref{conj:modular}) reframes the barrier as a function of height in the upper half-plane. This connects the problem to classical automorphic forms without invoking the heat kernel, which the numerical evidence in Section~\ref{sec:post-barrier} shows is structurally incompatible with $B(L)$.

\paragraph{What the explorations ruled out.} Four natural candidates for making positivity automatic---the heat kernel trace on $\mathrm{SL}(2,\mathbb{Z}) \backslash \mathbb{H}$, a reproducing kernel Hilbert space structure, and Rankin--Selberg L-values of $\Delta$---all fail for structural reasons documented in Section~\ref{sec:post-barrier}. Combined with the Cram\'er control, these negative results sharpen the problem: any proof of $B(L) > 0$ must use second-order information about the prime distribution, and the positivity must emerge from the quadratic form $Q_W$ itself rather than from an embedding into an a~priori positive structure.

\medskip
\noindent A companion paper~\cite{VaughanQuat} extends these results to quaternionic space, showing that $\pi$ and the finite primes separate along orthogonal quaternionic directions with a coherence ratio exceeding $1000{:}1$, and reformulating the barrier in adelic terms where positivity is automatic.

\bibliographystyle{plain}
\begin{thebibliography}{99}
\bibitem{CCM2025} A.~Connes, C.~Consani, H.~Moscovici, \emph{Zeta Spectral Triples}, arXiv:2511.22755 (2025).
\bibitem{CC2020} A.~Connes, C.~Consani, \emph{Weil positivity and Trace formula, the archimedean place}, Selecta Math.\ \textbf{27}, 27 (2021); arXiv:2006.13771.
\bibitem{CM2022} A.~Connes, H.~Moscovici, \emph{The UV prolate spectrum matches the zeros of zeta}, PNAS \textbf{119}(22), e2123174119 (2022).
\bibitem{Connes2026} A.~Connes, \emph{The Riemann Hypothesis: Past, Present and a Letter Through Time}, arXiv:2602.04022 (2026).
\bibitem{VaughanQuat} J.T.~Vaughan, \emph{Quaternionic structure of the completed Riemann zeta function: the archimedean--finite separation and the adelic barrier}, preprint, April 2026.
\end{thebibliography}

\end{document}
